---
title: "Mental Health and COVID-19"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: spacelab
---

```{r, include = FALSE}
library(tidyverse)
library(modelr)
library(p8105.datasets)
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

By: Armaan Ahmed, Brandon Rojas, Hansheng Zheng, Yueli Ma 

https://albinomeese.github.io/hz2711.github.io/
https://brojasnyc.github.io/br2598.github.io/
https://armaanahmed.github.io/

## Motivation

“It would be a tragedy if we beat back one public health crisis only to allow another to grow in its place.”


The COVID-19 pandemic has left an unprecedented impact on mental health, with especially staggering effects on the youth, elderly, and vulnerable populations, the magnitude of which was captured by multiple institutions in the United States including the U.S. Census Bureau. It has reached a point that earlier this week, U.S. Surgeon General Vivek Murthy published a 53-page advisory reviewing circumstances behind a decade long increase and outlining direct actions institutions including state and federal governments can take to combat it. 

There are many possible contributing factors to the increase in mental health disorders connected to the pandemic, including emotional hardship from deaths, a worsening opioid crisis, income volatility, and contracting the disease itself, along with lifestyle changes that came with the national effort against the disease. 

Early in the pandemic, lockdowns were enacted to encourage Americans to stay home and slow the spread of COVID-19. Complications of the lockdown included mass unemployment, widespread shutdown of small businesses, including an <a href="https://www.nytimes.com/2020/08/03/nyregion/nyc-small-businesses-closing-coronavirus.html">estimated one third of small businesses in NYC</a>, a <a href="https://www.bloomberg.com/news/articles/2020-06-08/black-business-owners-suffer-41-drop-in-covid-19-lockdowns">decline of 41% of black business owners</a> (If paywalled, <a href="https://drive.google.com/file/d/1so5hfe7FaEf6Newl-XENT8gfIQWiQKb9/view?usp=sharing">read here</a>), and increased incidence of mental health issues <a
href="https://icap.columbia.edu/tools_resources/silver-survey-summary-sheet/">in the elderly</a>. 


In addition, there is shortage of mental health professionals equipped to handle the increased incidence of cases. 


Innovations, whether from government or industry are needed and we will examine why. 


## Related Work


https://www.hhs.gov/sites/default/files/surgeon-general-youth-mental-health-advisory.pdf 

We compare the results with 2019 values from the National Health Interview Survey (NHIS) <a href="https://www.cdc.gov/nchs/data/nhis/mental-health-monthly-508.pdf/"> "Estimates of Mental Health Symptomatology" </a>

## Initial questions


•	What age groups have had the highest increase in anxiety or depression? (Which groups were most affected?) Repeat for race/etc

•	So on...
    

## Data

In April 2020, one month after COVID-19 was declared a national emergency, the National Center for Health Statistics (NCHS) teamed up with the Census Bureau to create the Household Pulse Survey, a 20 minute survey intended to record the impact of the pandemic on mental health.

For cleaning, the data required all dates be converted to Lubridate "Date" types from the original "Character", and columns converted to factors as needed.

COVID mortality data was filtered by date and scaled as appropriate.

We will examine the first two subsets of data focusing on incidence of anxiety and depression, and mental healthcare during the start of the pandemic.


Anxiety and Depression - NCHS


Where possible, we compare the results with 2019 values from the National Health Interview Survey (NHIS) as a baseline. <a href="https://www.cdc.gov/nchs/data/nhis/mental-health-monthly-508.pdf/"> "Estimates of Mental Health Symptomatology" </a>

https://data.cdc.gov/NCHS/Indicators-of-Anxiety-or-Depression-Based-on-Repor/8pt5-q6wp



## Anxiety and Depression - NCHS - Household Pulse Survey

The "Value" column represents the percentage of members of the subgroup column that experienced either symptoms of a depressive or anxious disorder, or both.

The value score is derived from a modified combination of two screening tests, the two-item <a href = "https://www.hiv.uw.edu/page/mental-health-screening/phq-2/"> </a> Patient Health Questionnaire (PHQ-2) and the two-item <a href = "https://www.hiv.uw.edu/page/mental-health-screening/gad-2 /"> Generalized Anxiety Disorder (GAD-2)</a>  scale. 

```{r}
anxiety_depression_df = read.csv("Indicators_of_Anxiety_or_Depression_Based_on_Reported_Frequency_of_Symptoms_During_Last_7_Days.csv") %>% 
  janitor::clean_names() %>% 
   na.omit

head_ad = anxiety_depression_df %>% select(i_indicator, state, subgroup, time_period_label,value) %>% knitr::kable() 

head(head_ad)
```
## Mental Health Care - NCHS - Household Pulse Survey

Four months into the survey, questions were added about the impact of the pandemic on mental healthcare. Estimates are derived from the same survey and the "value" column represents the percentage of members of the subgroup column who done one of the following:

a) took prescription medication for their mental health

b) received counseling or therapy from a mental health professional 

c) needed counseling or therapy from a mental health professional but did not get it (unmet mental health need)


```{r}
mh_care_df = read.csv("Mental_Health_Care_in_the_Last_4_Weeks.csv") %>%  
  janitor::clean_names() 

head_mh = mh_care_df %>% 
  select(i_indicator, state, subgroup, time_period_label,value) %>% 
  knitr::kable() 

head(head_mh)

```


## Discussion:
 

## What we can do:


Everyone reading this can make a difference, if you’re at Mailman you have, or will have, the skills to leave an impact on the future of mental health. 

In the wake of an <a href="https://usafacts.org/articles/over-one-third-of-americans-live-in-areas-lacking-mental-health-professionals/">intractable shortage of mental health professionals</a>, universal or accessibly priced, mobile, around the clock access to mental health treatment will be paramount for society to tackle the sharp increase in mental health disorders in the past two years.

There are immense career opportunities in the field of mental health, including telemedicine, behavioral healthcare, online therapy, and accessibility innovations using brain computer interfaces.

Since 2020, <a href="https://www.forbes.com/sites/katiejennings/2021/06/07/venture-funding-for-mental-health-startups-hits-record-high-as-anxiety-depression-skyrocket/?sh=4c4ea68a1116">record numbers of mental health startups have received venture capital funding</a>. 

A few examples:

• Online mental health/wellness service <a href="https://www.fiercehealthcare.com/tech/softbank-leads-mental-health-startup-cerebral-s-300m-round-propelling-valuation-to-4-8b.%20Published%20December%208,%202021.">Cerebral receiving $300M</a> from in a series C round led by SoftBank to expand services on its 200,000 patients

• <a href="https://www.fiercebiotech.com/medtech/digital-therapist-woebot-perks-up-90m-for-ai-powered-mental-health-platform.%20Published%20July%2021,%202021.%20Accessed%20December%2010,%202021.">Woebot closes $90M</a>: A free “AI/NLP powered chatbot” app that uses your personal input and generates individualized treatment using cognitive behavioral therapy inspired techniques.</p>

• <a href="https://www.whatif.vc/portfolio">And more</a>