---
title: "Access to Mental Health"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: spacelab
    code_folding: hide
---

```{r, include = FALSE}
library(tidyverse)
library(modelr)
library(plotly)
library(lubridate)
library(dplyr)
library(readr)
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

`%notin%` <- Negate(`%in%`)
```

## Mental Healthcare and the Pandemic

Using plotly we illustrate the increase of mental health burden in the US by age group.

We include a crude graph of COVID mortality from Oct 2020 to Oct 2021 to add context

```{r}
cm_df = read_csv("Provisional_COVID-19_Death_Counts_by_Week_Ending_Date_and_State.csv") %>% 
  janitor::clean_names() %>%
  mutate(
    start_date = as_date(start_date, format = '%m/%d/%Y'), 
    end_date = as_date(end_date, format = '%m/%d/%Y'),
   ) %>%
   filter(end_date >=("2020-08-31") & end_date <= ("2021-10-11")) %>%
  filter(state == "United States") %>%
  filter(group == "By Week")

cm_anxdep_df = read_csv("Provisional_COVID-19_Death_Counts_by_Week_Ending_Date_and_State.csv") %>%  #Anxiety survey began earlier
  janitor::clean_names() %>%
  mutate(
    start_date = as_date(start_date, format = '%m/%d/%Y'), 
    end_date = as_date(end_date, format = '%m/%d/%Y'),
   ) %>%
   filter(end_date >=("2020-05-05") & end_date <= ("2021-10-11")) %>%
  filter(state == "United States") %>%
  filter(group == "By Week")

z = cm_df %>% 
  ggplot(aes(x = end_date, y = covid_19_deaths)) + 
  geom_line()
ggplotly(z)
```

Cleaning Mental Healthcare data, filtering separately by age and state
```{r}
mh_care_df = read_csv("Mental_Health_Care_in_the_Last_4_Weeks.csv") %>%  
  janitor::clean_names() %>% 
  mutate(
    time_period_start_date = as_date(time_period_start_date, format = '%m/%d/%Y'), 
    time_period_end_date = as_date(time_period_end_date, format = '%m/%d/%Y'),
    state = as_factor(state),
    subgroup = as_factor(subgroup),
    group = as_factor(group),
    indicator = as_factor(indicator)    
   )

mh_age = c("18 - 29 years", "30 - 39 years", "40 - 49 years", "50 - 59 years", "60 - 69 years", "70 - 79 years", "80 years and above")

mh_age_df = mh_care_df %>% 
  filter(subgroup %in% mh_age) %>% 
  #filter(indicator %in% "Took Prescription Medication for Mental Health And/Or Received Counseling or Therapy, Last 4 Weeks")
  #filter(indicator %in% "Took Prescription Medication for Mental Health, Last 4 Weeks")
  filter(indicator %in% "Needed Counseling or Therapy But Did Not Get It, Last 4 Weeks")
  #filter(indicator %in% "Received Counseling or Therapy, Last 4 Weeks")

mh_state = c("By Race/Hispanic ethnicity", "United States", "By Disability status", "By Gender identity", "By Age", "By Sex", "By Education", "National Estimate", "By Presence of Symptoms of Anxiety/Depression", "By Sexual orientation")

mh_state_df = mh_care_df %>% 
  filter(group %notin% mh_state) %>% 
  #filter(indicator %in% "Took Prescription Medication for Mental Health And/Or Received Counseling or Therapy, Last 4 Weeks")
  #filter(indicator %in% "Took Prescription Medication for Mental Health, Last 4 Weeks")
  filter(indicator %in% "Needed Counseling or Therapy But Did Not Get It, Last 4 Weeks")
  #filter(indicator %in% "Received Counseling or Therapy, Last 4 Weeks")

mh_drug_care_age_df = mh_care_df %>% 
  filter(subgroup %in% mh_age) %>% 
  filter(indicator %in% "Took Prescription Medication for Mental Health And/Or Received Counseling or Therapy, Last 4 Weeks")
  #filter(indicator %in% "Took Prescription Medication for Mental Health, Last 4 Weeks")
  #filter(indicator %in% "Needed Counseling or Therapy But Did Not Get It, Last 4 Weeks")
  #filter(indicator %in% "Received Counseling or Therapy, Last 4 Weeks")

```

Cleaning Anxiety Data
```{r}
anxdep_df = read_csv("Indicators_of_Anxiety_or_Depression_Based_on_Reported_Frequency_of_Symptoms_During_Last_7_Days.csv") %>%  
  janitor::clean_names() %>% 
  mutate(
    time_period_start_date = as_date(time_period_start_date, format = '%m/%d/%Y'), 
    time_period_end_date = as_date(time_period_end_date, format = '%m/%d/%Y'),
    state = as_factor(state),
    subgroup = as_factor(subgroup),
    group = as_factor(group),
    indicator = as_factor(indicator)
   )

anxdep_age = c("18 - 29 years", "30 - 39 years", "40 - 49 years", "50 - 59 years", "60 - 69 years", "70 - 79 years", "80 years and above")

anxdep_age_df = anxdep_df %>% filter(subgroup %in% mh_age) %>% 
  #filter(indicator %in% "Symptoms of Anxiety Disorder")
  filter(indicator %in% "Symptoms of Anxiety Disorder or Depressive Disorder")
  #filter(indicator %in% "Symptoms of Depressive Disorder")
  
anxdep_state = c("By Race/Hispanic ethnicity", "United States", "By Disability status", "By Gender identity", "By Age", "By Sex", "By Education", "National Estimate", "By Presence of Symptoms of Anxiety/Depression", "By Sexual orientation")

anxdep_state_df = anxdep_df %>% filter(group %notin% anxdep_state) %>% 
  #filter(indicator %in% "Symptoms of Anxiety Disorder")
  filter(indicator %in% "Symptoms of Anxiety Disorder or Depressive Disorder")
  #filter(indicator %in% "Symptoms of Depressive Disorder")
```


```{r}
#Date vs subgroup w/ facet w/ CI
# mh_age_df %>%
#   ggplot(aes(x = time_period_start_date, y = value, color = subgroup)) + 
#   geom_line() + 
#   geom_ribbon(aes(ymin=low_ci, ymax=high_ci), linetype=2, alpha=0.1)  + 
#   facet_wrap(vars(mh_age_df$subgroup)) +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1), axis.text.y = element_text(angle = 45, hjust = 1)) +
#   scale_y_continuous(breaks = seq(0, 40, 5)) +
#   labs(x = "Month", y = "% Subgroup affected")
# 
# #Date vs subgroup, no facets w/ CI
# mh_age_df %>%
#   ggplot(aes(x = time_period_start_date, y = value, color = subgroup)) + 
#   geom_line() + 
#   geom_ribbon(aes(ymin=low_ci, ymax=high_ci), linetype=2, alpha=0.1)  + 
# #  facet_wrap(vars(mh_age_df$subgroup)) +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1), axis.text.y = element_text(angle = 45, hjust = 1)) +
#   labs(x = "Month", y = "% Subgroup affected", title = "Healthcare inaccessibility throughout the pandemic by age group ")
```

## Mental Health Burden by Age Group

The following is presented alongside COVID mortalities scaled by tens of thousands to give context on the state of the pandemic at time of recording. 

We notice that across age groups, there is a uniform increase in the use of mental health resources, particularly in the 18-29 cohort.

```{r}
# mental health availability w/ facet w/o CI w/COVID mort
g = ggplot() + 
  geom_line(data = mh_age_df, aes(x = time_period_start_date, y = value, color = subgroup) ) +
  geom_line(data = cm_df, aes(x = end_date, y = covid_19_deaths * 0.001) ) + 
  #geom_ribbon(aes(data = mh_age_df, ymin=low_ci, ymax=high_ci), linetype=2, alpha=0.1)  +
  facet_wrap(vars(subgroup)) +
  scale_y_continuous(sec.axis = sec_axis(trans = ~.*001)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  labs(x = "Month", y = "% without access to mental health treatment",title = "Proportion of those unable to find mental health care when needed")
ggplotly(g)

y = ggplot() + 
  geom_line(data = mh_drug_care_age_df, aes(x = time_period_start_date, y = value, color = subgroup) ) +
  geom_line(data = cm_df, aes(x = end_date, y = covid_19_deaths * 0.001) ) + 
  #geom_ribbon(aes(data = mh_age_df, ymin=low_ci, ymax=high_ci), linetype=2, alpha=0.1)  +
  facet_wrap(vars(subgroup)) +
  scale_y_continuous(sec.axis = sec_axis(trans = ~.*001)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  labs(x = "Month", y = "% using therapy, psychiatric drugs", title = "Use of therapy or psychiatric drugs by age group")
ggplotly(y)

## % Screened for Anxiety or Depression by age
x = ggplot() + 
  geom_line(data = anxdep_age_df, aes(x = time_period_start_date, y = value, color = subgroup) ) +
  geom_line(data = cm_anxdep_df, aes(x = end_date, y = covid_19_deaths * 0.001) ) + 
  #geom_ribbon(aes(data = mh_age_df, ymin=low_ci, ymax=high_ci), linetype=2, alpha=0.1)  +
  facet_wrap(vars(subgroup)) +
  scale_y_continuous(breaks = seq(0, 70, 20), sec.axis = sec_axis(trans = ~.*001)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  labs(x = "Month", y = "% Subgroup affected", title = "% Positive screening for Anxiety or Depression by age")

ggplotly(x)


```


## Hypothesis Test

We will explore whether there may be differences in the proportion of individuals unable to find mental healthcare.

We use one-way ANOVA because we are comparing within a single categorical variable with more than two levels.

$H_0$ : There is no difference in accessibility of mental healthcare by states

$H_1$ : At least two states have unequal access for healthcare

```{r}
state_means = mh_state_df %>% 
  select(subgroup, value)%>%
  group_by(subgroup) %>% 
  na.omit() %>%
  summarize(mean = mean(value),
            se = sd(value)
            )
  knitr::kable(head(state_means,caption = "Average % incidence of anxiety/depression by State during time period"))

```

Based on the below we conclude with beyond 99.99% confidence that there is unequal access between states and the accessibility of mental healthcare. 
```{r}
anova_state_mh_access = aov(value ~factor(subgroup), data = mh_state_df)
summary(anova_state_mh_access) 
```

We do a similar test to see if there are differences between age group and access to care, and reach a similar conclusion:
```{r}
anova_age_mh_access = aov(value ~factor(subgroup), data = mh_age_df)
summary(anova_age_mh_access) 
```


And again, same results for for positive screening for anxiety or depression by state:

```{r}
anova_state_ax = aov(value ~factor(subgroup), data = anxdep_state_df)
summary(anova_state_ax) 
```