---
title: "P8105 Final Project Project Proposal"
author: "Armaan Ahmed, add your names "
date: "11/18/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Define your teams and propose a project. This proposal should be a half-page to a page in length and include:


The group members (names and UNIs)
Armaan Ahmed - asa2249
Hansheng Zhang - hz2711
Yueli Ma - ym2814’
Brandon Rojas br2598


The tentative project title


The motivation for this project


The intended final products


The anticipated data sources
https://data.cdc.gov/NCHS/Indicators-of-Anxiety-or-Depression-Based-on-Repor/8pt5-q6wp


The planned analyses / visualizations / coding challenges


The planned timeline

There should be one proposal per group, written collaboratively using .Rmd (rendering to a GH document) in a dedicated GitHub repo. This is intended to describe a project that is in the planning stages. Only one team member should submit this proposal via Canvas. 
Please submit your GitHub repo containing your team's proposal here.


```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
